### 2017-12-06 doc/MSE_OCT.tex
last model version

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
\usepackage{array,multirow,graphicx}
\usepackage{pdflscape}% for landscape

\usepackage{adjustbox}     % adjusting table (too wide)
\usepackage{rotating}      % sidewaytables
\usepackage{array}
\usepackage{stackengine}

\usepackage{amsmath} % \numberwithin{equation} doesn't exist without this package.
\numberwithin{equation}{subsection} % This line resets equation numbering when starting a new section.
\renewcommand{\theequation}{Eq. \thesubsection.\arabic{equation}} % This line ads "Eq." in front of your equation numbering.

\begin{document}

\title{MSE and dynamic state variable model with sigma function}

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Jan Jaap Poos}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{janjaap.poos@wur.nl}}

\abstract{
TBD
Farcas 2016:The case of fisheries management illustrates how the inherent structural instability of ecosystems can have deep-running policy implications. We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses complex food webs in its operating model. 
Use DSVM with a tuning parameter that measures how important is to be near optimal.
We use this DSVM to investigate how fishing locations change in response when fish biology is incorporated.}

\date{\today}

\keywords{word1; word2; word3; and word4}

\maketitle

\section{Introduction}
\label{sec1}

The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices. Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other. LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation.

Fisheries management typically focuses on the sustainable exploitation of single target species, but the potentially negative effects on, for example, reaching other species conservation objectives, should also be evaluated. Achieving single species MSY in complex and dynamic fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017}. There is no unique way of translating the single-species MSY objective to the multispecies case. Maximisation of yield from one stock will generally require different strategies than maximisation of yield from another.  Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures. Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty.

MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}. At the same time being simple enough to build and calibrate reliably with the available data. More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.

There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010} have been developed. These models generally predict the effect in the short term (within a year) by optimizing an objective function and determine which area best suits this behaviour given the set of incentives that exist (which may also depend on factors such as size, home port, distance to fishing grounds and expected catch rates). Effects of e.g. relocation costs of marine protected areas, the implementation of the landings obligation, have been modelled under these approaches. In contrast, to enable longer term prediction accounting for the impact of these choices on the populations and consequently on the future yields, this fleet dynamic model should be coupled to a biological dynamic model where the feedback between fleet and stock dynamics are explicity modelled.

We review the potential of MSE, emphasazing the role of individual harvester decision-making and socioeconomic drivers on management effectiveness, i.e. if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}. Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}. The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel. The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.

Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.

\section{Methods}
\label{sec2}
The MSE approach is based upon a set of model of the "true" population dynamics of the species (called "operating model"; section \ref{sec2.1}). The operating model aims capturing the key processes in the dynamics of the fish population given the best knowledge available, and can be thought of as a minimum realistic model \cite{Punt1995}. The next step in the MSE is to simulate the process of monitoring the stock (called "observation model"; section \ref{sec2.2}). This resulted in simulated measurements such as individual harvester decision-making and the consequent biomass (including error). The observation data are then passed to the "management model" (section \ref{sec2.3}). The management model encompasses the harvest control rules (HCR). The HCR references to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level, changes in gear or spatial and temporal reallocations or restrictions of fishing effort. Management actions are rarely implemented without error. This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g. when and where harvest is taken) are not accounted for in the HCR. By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.

\subsection{Stock structure and dynamics}
\label{sec2.1}

For purposes of model simplicity, we considered a hypothetical mixed fishery on two species. The fishery was modelled in a individual based dynamic state variable model. In order to allowed for fishing effort allocation, the model was seasonally and spatially explicit. The dynamics of the two fish species were governed by a simple age-structured model that is also seasonally and spatially explicit. In the age structured model for the fish species, individuals were assumed to be born at age 0, and in season 1. The number of fish of age $a$ at year $y$, in season $s$ and area $p$ was written as $N (a, y, s, p)$. The number of new-born individuals in the model was independent of the size of the adult population. Individuals could be born in each of the areas, such that  
\begin{equation}
N (0, y, 1, p) = R (p)
\end{equation}

Mortality in the model could occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens. However, for simplicity we assumed that his natural mortality was negligible. Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species. The decrease in population numbers was thus the result of the catches ($C (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area. 
\begin{equation}
C (a, y, s, p) = q(a) * E(y,s,p),
\end{equation}
and 
\begin{equation}
N (a, y, s, p) = N (a, y, s, p) - C (a, y, s, p). 
\end{equation}

It was assumed that fishing mortality ocurred before migration ocurred. Migration for each species was defined by a matrix $Mig$ that defines the impact of migration (.e. immigration-emigration) on a given stock was also dependent on the relative sizes of the stocks. The size of that matrix  was defined by the number of age classes, seasons and number of immigrants into area $p$ ($im$) and emigrants leaving area $p$ ($em$) in the model. The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($p (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs.   
\begin{equation}
Mig (a, 1, s, im, em) = Mig(a, 1, s, im, em) + N (a, y, s, em) * p (a, y, s, i, em, im),
\end{equation}
and then population numbers
\begin{equation}
N (a, y, s, im) = N (a, y, s, im) +  Mig (a, 1, s, im, em).
\end{equation}

Life history traits inputted were: (i)  during the first season, specie 1 spawned in $A$, while specie 2 spawned in $B$, (ii) every season, 20\% of specie 1 and 2 emigrated and immigrated between areas, (iii) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (iv) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (v) at the beginning of each year the fishing recruitment was stablish in 100.

The purpose of the operating model was not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.

\subsection{The dynamic-state variable model}
\label{sec2.2}

To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}. A dynamic state variable model \cite{Houston1999, ClarkandMangel2000} was used to model location choice, extending the model structure in \cite{Batsleer2015}. In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations. The model had annual fines for exceeding landings or catch quota as in \cite{Alzorriz2018}). In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings or catch quotas at the end of the year:
\begin{equation}
\Phi (C_i, Q_i, F_i)= -\sum_i (max( 0, (C_i - Q_i))* F_i),
\end{equation}
where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel. These cumulative landings defined the state of the individual. $Q_i$ was the annual individual quota for landings or catches for the quota species. Individual quotas were not transferable. In order to reduce computation time, we set quota only for two species. $F_i$ was the fine per unit weight for exceeding individual landings or catch quotas was set to $3 x 10^3$ Euro day\textsuperscript{-1}. These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.

The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, s)= \Phi (C_i, Q_i, F_i)$. For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$ (A, B and including area 0: “staying in port”), and to keep or discard any combination of the age classes caught of the quota species. This discarding was defined by a matrix $d$. The size of that matrix was defined by the number of species under quota constraints and the number of age classes. Each element could take the value 0 (discard) or 1 (keep on board and land). The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
\begin{equation}
V (C_i, Q_i, F_i, s) = max_{p,d}( R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1])),
\end{equation}
where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependant over fish age and time (subsection \ref{sec2.2.1}). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table ~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices. The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

Rather than assuming that each individual always made the optimal choice, we assigned a probability to each choice proportional to its expected utility, following \cite{Dowling2011}. The expected utility for any choice was
\begin{equation}
U (C_i, Q_i, F_i, s) = R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]).
\end{equation}

If $U^*$ was the expected utility at the optimal choice for a given t, we set
\begin{equation}
\Delta_{p, d}(C_i, Q_i, F_i, s) =  U^* (C_i, Q_i, F_i, s) - U (C_i, Q_i, F_i, s),
\end{equation}
and then defined the probability of a choice for a given area and discarding as	
\begin{equation} \label{eq:solve}
P_{p, d}(C_i, Q_i, F_i, s) = \frac
                % Nominator
                {e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}}
                % Denominator
                {\sum_p \sum_d e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}},
\end{equation}
where $\sigma$ was a tuning parameter that measures how important was to be near optimal. A high value of sigma resulted in uniform probabilities of choices, and that vessels were distributed uniformly among the different fishing areas. In contrast, if sigma was small then all vessels concentrated in the optimal location (but note that $\sigma$ should be $> 0$). For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 1.1 x 10^{11}$, but was generally of the order of $1.1 x 10^5$ in magnitude).

The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location and discarding behaviour for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota. Further details for this procedure can be found in \cite{Alzorriz2018, Batsleer2016} and \cite{Dowling2011}.

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in ~\ref{eq:solve}. This was done for 700 individual vessels.

\begin{table}[!h]
\centering
\caption{Fishery parameterization.}
\label{t:fishery}
\resizebox{0.7\textwidth}{!}{%
\begin{tabular}{@{}lc lc @{}} 
\hline
\textbf{model}                              &               & \textbf{Variable costs}                        &      \\\hline%&\textbf{Market value}&  \\ \hline
& & & \\%& & \\
number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  \\%& sp1 (Euro kg\textsuperscript{-1})                 & 1000 \\ 
$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0   \\% & sp2 (Euro kg\textsuperscript{-1})                 & 1000   \\
number of areas                             & 2             &crew share                                      & 35\% \\%& & \\
number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    \\%& &\\
intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
& & & \\\hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependant pricing}
\label{sec2.2.1}
% Prices of fish have been assumed to be constant over fish age and time.
Common knowledge and case studies indicated that size and weight of a fish could be important attributes when determining its economic value. However, so far there has been no broader assessment whether the size-dependence of price really is prevalent and significant in fisheries. The results from \cite{Zimmermann2011, Zimmermann2013} underline that size-dependent pricing is common in Norwegian fisheries, and is therefore also of high relevance for decisions in fisheries management. Therefore, the resulting data of mean weights, average prices and total yield per weight classes and year were used for further analysis. We fitted linear mixed models (Bates and Maechler, 2010) and general additive models (Wood, 2010) in R (R: Development Core Team, 2012), using mean price and mean body weight as the response variable and covariate, respectively. Thus, the statistical models took the form  
\begin{equation}
 p_{i} \sim c_{0} + c_{1} w_{i} + f (y_{i}), 
\end{equation}
where p is price of observation i, wi is the mean weight of the corresponding weight class, and the term f(yi) corresponds to the effect of year, treated either as a random effect or as a smooth term. Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models. Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results. All models were fitted independently to each stock.
\begin{table}[!h]
\centering
\caption{Size dependant pricing.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}lcccccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{6}{c}{season} \\ 
age                      & 1     & 2     & 3    & 4    & 5    & 6    \\ \hline
                      \\
1                       &    &    &   &   & & \\
2                       &    &    &   &   & &\\
3                       &    &    &   &   & &\\
4                       &    &    &   &   & &\\
\\
\hline 
\end{tabular}
}
\end{table}

\subsection{Endogenously determined quotas}
\label{sec2.3}

The three management strategies we considered differ by the formulae used to determine exploitation rates (HCR). These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan. For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.

\begin{table}[!h]
\centering
\caption{}
\label{scenarios}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c  c }
\hline
                                & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
                                & \textit{(landings selectivity)}  & \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
& & &  \\
Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
Manager catch perception        & $C = L + D$                    &        $C = L$                       &  $C = L$  \\
Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\
Selectivity ($L/C$)             &  $ \neq 1$                     & $=1$                                 & $=1$\\
&  & &  \\\hline
\end{tabular}
}
\end{table}

\begin{enumerate}

\item
\textbf{Scenario I: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield. Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.

\item
\textbf{Scenario II: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.

\item 
\textbf{Scenario III: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding. However the manager could think that is fully enforced and assumed that only what is landed contribute to yield. This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality. Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.

\end{enumerate}

HCRs in the form of Quotas were determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeded Fmsy. This required that fishing mortality for each stock must not exceed Fmsy, which was a limit reference point. An Optimum Yield (OY) was determined based on a MSY that must prevented overfishing, and that could be reduced from the MSY limit to take into account social, economic and precautionary considerations.  Therefore, quota was determined by a set of harvest control rule parameters consisting of the exploitaion rate at MSY.

The resulted biomass harvested rate in a year $y$, depended in the manager catch perception and the standing biomass:
\begin{equation}
 H_{a, y, s} =\frac
                % Nominator
                {C_i numbers(a, y, s, p)}
                % Denominator
                {C_i numbers(a, y, s, p) \times B_i numbers(a, y, s, p)},
\end{equation}

\begin{equation}
 lratio (a, y, s) = \frac
                % Nominator
                {L_i numbers(a, y, s, p)}
                % Denominator
                {C_i numbers(a, y, s, p)},
\end{equation}

\begin{equation}
 H_{MSY} = YIELD CURVE..., 
\end{equation}

\begin{equation}
 Q_{MSY} = \frac
                % Nominator
                {\sum( [\frac{H_{MSY}}{\bar{H}} \times H (a, s) \times lratio (a, s) \times B numbers(a, s)] \times wts) }
                % Denominator
                {number of vessels}, 
\end{equation}
\begin{equation}
 Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85)), 
\end{equation}

%Given the landing parameters associated

%hr<- catches.n/ (catches.n+ pop)
%landings.ratio <- landings.n/ cathces.n
%OY <- 
%hr_MSY <- OY[OY_landings==max(OY_landingslandings),]_hr

%prel.quota <-  sum(sweep((hr1wanted[1]/mean(hr1[,yy,]))* hr1[,yy,]*landings.ratio1[,yy,]*apply(pop1[,yy+1,,],c(1,2), sum) ,1,wts,"*"))/SIMNUMBER
%        tac.constrained <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
%        quota1[,yy+1,,] <- max(min(prel.quota, tac.constrained[2]), tac.constrained[1])

%From an economic perspective, a linear increase towards maximum TAC is the optimal strategy under plausible assumptions on market demand for fish and harvesting cost, as shown in Froese et al., 2010. The economic rationale is that, beyond Blim, the stock is within safe biological limits, and as it becomes more and more productive, it is able to supply the market to an increasing extent.

%A linear reduction in fishing mortality has been implemented in Australia (DAFF 2007) a nd is common in the USA. The differences between a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in Appendix S4.

%estimating the hr Fisheries are managed by a total allowable catch (TAC). A maximum TAC is set for each stock so that the respective target biomass is maintained
%on average. This maximum TAC may be taken as long as biomass fluctuations remain above Bmsy.

%- No quota constrained

%- Through a yield per recruit model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.  Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield. The principal purpose of yield per recruit models is to get an idea of the effect of selection pattern and fishing mortality on the yield from a fixed number of individuals that enters the fishery.

%Therefore if we have a harvest rate of 0.32, we will need to reduce the harvest in a 68\%. Reduce quota in next run in 68\%. From this reduction in quota and the subsequent reduction in IQ, this scenario will be evaluated allowing and not allowing the fleet to discard.

\section{Results}
\label{sec3}

\newpage
%\begin{landscape}
Scenario I:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand.eps} 
\end{minipage}
\caption{}
\end{figure}
%\end{landscape}

%----------------------------------------------------------------------------------------------------------------------------------------

Scenario II:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO.eps} 
\end{minipage}
\caption{}
\end{figure}

%----------------------------------------------------------------------------------------------------------------------------------------

Scenario III:

\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_nocompliance.eps}  
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\centering
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_nocompliance.eps} 
\end{minipage}
\caption{}
\label{fig:effort_lo}
\end{figure}

%----------------------------------------------------------------------------------------------------------------------------------------

\subsection{Simulated catch decisions and location choice}
\label{sec3.1}

The largest yield (or catch) that can be taken from a species' stock over an indefinite period. Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely. Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small. At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate. At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals. Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity. At this point, there are no surplus individuals to be harvested and yield drops to zero. The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield

There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983). This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource. 
For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis. Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.

As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing. It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere. Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g. Panayotou 1988).

The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced. This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g. the tuna fishery for the canning industry.

The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition. Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points. Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.

In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points. In evaluating investment projects, including resource management, economists discount the future value of any commodity. Discount rates may be in the order of 10\%. In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested. Long-lived species with slow growth rates, such as whales, clearly fall into this category. The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).

\newpage
\section{Discussion}
\label{sec4}
tbd
prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative. Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future. This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.

\begin{notes}[Acknowledgements]
The content of this paper does not reflect the official opinion of the European Commission. Responsibility for the information and views expressed in this paper relies entirely with the authors.
\end{notes}

\newpage
\bibliographystyle{plain}
\bibliography{mylib}

\end{document}