### 2018-08-30 doc/MSE_MAR_18.tex
some discussion

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
\usepackage{array,multirow,graphicx}
\usepackage{pdflscape}% for landscape

\usepackage{adjustbox}     % adjusting table (too wide)
\usepackage{rotating}      % sidewaytables
\usepackage{array}
\usepackage{stackengine}

\usepackage{amsmath} % \numberwithin{equation} doesn't exist without this package.
\numberwithin{equation}{subsection} % This line resets equation numbering when starting a new section.
\renewcommand{\theequation}{Eq. \thesubsection.\arabic{equation}} % This line ads "Eq." in front of your equation numbering.

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\begin{document}

\title{MSE and dynamic state variable model with sigma function}

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Jan Jaap Poos}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{janjaap.poos@wur.nl}}

\abstract{Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY). The EU common fisheries policy, designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the MSY reference points as targets for exploiting commercially important fish stocks. To incentivise individual business to improve selectivity and avoid unwanted catches, it also introduced a landing obligation. Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests. The impact of exploiting at MSY and no discarding allowance on the optimal effort allocation between fleets is analyzed. We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model. The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations. Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation. Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.4 times larger profits than before the CFP implementation. The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}

\date{\today}

\keywords{management strategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield}

\maketitle

\section{Introduction}
%We have to make sure we are clear on the ages of the species: do we start at 0 or at 1?

Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulation\cite{Punt2016}. Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}. Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable. As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy. The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.

The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}. Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed. These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual. The effect of a choice on the utility depends on the economic environment, such as the home port of the vessel and the distance to fishing grounds, and the biological enviroment, such as the spatial distribution of the resources. Using such dynamic variable state models, effects of e.g. relocation costs of marine protected areas have been modelled \cite{Dowling2011}.

Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995}. These studies may shed light on the potential outcomes of the European fisheries management reform of 2013. That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}. Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO. These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}. This will result in fishing effort reallocation and early closures of fisheries once quota have been reached. However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013, Ulrich2017}. Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
This will result in fishing effort reallocation and early closures of fisheries once quota have been reached. However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. 
To enable longer term prediction of the effects of changes in fisheries management, and thus complete the MSE, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response. We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e. if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}. Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}. The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel. The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels. To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.

\section{Methods}

Managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species. In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota. Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold. Each of these elements is discussed in more detail below. The essential elements of the framework are summarized in Figure \ref{fig:MSE} and \ref{fig:stockdyn}.

The framework consists of an operating model (OM) and a management procedure (MP). The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.  The OM thus includes individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance (Fig. \ref{fig:stockdyn}). The management body used the MP to make its decision on how to respond to the state of the resource. This MP thus includes the data collection from the fishery, how these are interpreted, and the harvest control rule (HCR) that dictates the limits that are set to the fishery.

\begin{figure}[!h]
\centering
\includegraphics[width=0.55\textwidth]{Figures/MSE.eps} 
\caption{Conceptual overview of the MSE approach, including the OM and the MP components of the framework.}
\label{fig:MSE}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.4\textwidth]{Figures/Areadynamics.eps} 
\caption{The OM component, including the key processes in the dynamics of the fish populations.}
\label{fig:stockdyn}
\end{figure}

\subsection{Population dynamics}

To allow the fishery to make spatial and temporal choices, the model needed to be seasonally and spatially explicit. The dynamics of the fish stocks were modelled using an age-structured model that was spatially explicit, with seasonal time steps. In each seasonal time step, fish grow, migrate, and die. The number of fish of stock $i$ of age $a$ at year $y$, in season $s$, and area $p$ was written as $N_i (a, y, s, p)$.  The ages in the model range between age 0 and age $A$, the maximum age in the model. The seasons range between season 1 and season $S$, the last season within each year. Individuals were born at age 0, at the start of each year $y$, in season 1. The number of recruits $R_i (p)$ in the model was a function of the area $p$ and independent of the size of the adult population. The population numbers at age 0 are thus

\begin{equation}
N_i (0, y, 1, p) = R_i (p).
\end{equation}

Mortality in the model resulted from fishery catches and natural causes such as predation, diseases, and scenesens. The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$) and a natural mortality constant $M_i$ that described natural mortality as a fixed fraction of the population. These mortalities reduced the population numbers among a cohort of fish. Because the model was seasonally structured, the population numbers for seasons 2 to $S$ were dependent on the previous season,

\begin{equation}
N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p) - M_i N_i (a, y, s, p) . 
\end{equation}

Likewise, the population numbers in season 1 depended on the numbers in season $S$ of the previous year,

\begin{equation}
N_i (a+1, y, 1, p) = N_i (a, y, S, p) - C_i (a, y, S, p) - M_i N_i (a, y, s, p). 
\end{equation}

Migration for each species was defined by an array $D_i (a, s, em, im)$ that defined  immigration and emigration on a given stock relative to the stock sizes. The size of that array was defined by the number of age classes, seasons and number of areas. emigrants leave area $em$ and move to area $im$. The emigrated part of the population is then substracted from each of the areas, so that that population numbers per year and season remain unaffected by migration.

Individual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight. The weights for individuals in the stock and in the catches are thus calculated as

\begin{equation}
w_i(a,s) =  \alpha * ( L_{\infty_i} * (1-\exp^{(-K * (a+(s/S)))}))^{\beta}.
\end{equation}

The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model. The Dynamic State Variable model inputs consist of the expected individual catch rates, which are random variables. These random variables were normally distributed, with means $\hat c_i (a, y, s, p)$ being a function of population size, age dependent catchability $q_i (a)$ in any year, season, and area,

\begin{equation}
\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a)  * w_i(a,s).
\end{equation}
The standard deviations $\Sigma_i (a, y, s, p)$ of the catch distributions were constant fraction  their means, using a ratio $\eta$.

\subsection{Fleet dynamics}

To simulate a fleet of individual fishing vessels,we used a Dynamic State Variable Model \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}. The model was used to model location choice, same model structure as in \cite{Alzorriz2018}, but without the choice to discard one or more species and size classes while fishing. Each individual vessels had a set of choices, which include the choice to go fishing in a season, location choice within that season. The model had annual fines for exceeding landings quota as in \cite{Alzorriz2018}. In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings quotas at the end of the year:
\begin{equation}
\Phi (C_i, Q_i, F_i)= -\sum_i (\textrm{max}( 0, (C_i - Q_i))* F_i),
\end{equation}

where $C_i$ was the cumulative annual catches for species $i$ for an individual vessel. These cumulative catches defined the state of the individual. $Q_i$ was the annual individual quota for catches for the different species. Individual quotas were not transferable. $F_i$ was the fine per unit weight for exceeding individual catches quota.

The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, S)= \Phi (C_i, Q_i, F_i)$. For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$, or to stay in port. While fishing, any combination of the age classes caught of the quota species had to be landed. The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
\begin{equation}
V (C_i, Q_i, F_i, s) = max_{p}( R(p, s)- G(p) - C(p) + E_{p}[V (C_i^\prime, Q_i, F_i, s+1)]),
\end{equation}

where $R(p, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependent over fish age (subsection size dependent pricing). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices. The term $E_{p}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

Rather than assuming that each individual always made the optimal choice, we assigned a probability to each choice proportional to its expected utility, following \cite{Dowling2011}. The expected utility for any choice was
\begin{equation}
U (C_i, Q_i, F_i, s) = R(p, s)- G(p) - C(p) + E_{p}[V (C_i^\prime, Q_i, F_i, s+1)].
\end{equation}

If $U^*$ was the expected utility at the optimal choice for a given t, we set

\begin{equation}
\Delta_{p}(C_i, Q_i, F_i, s) =  U^* (C_i, Q_i, F_i, s) - U (C_i, Q_i, F_i, s),
\end{equation}

and then defined the probability of a choice for a given area and discarding as

\begin{equation} \label{eq:solve}
P_{p}(C_i, Q_i, F_i, s) = \frac
                % Nominator
                {e^{ -\Delta_{p}(C_i, Q_i, F_i, s)/\sigma}}
                % Denominator
                {\sum_p e^{ -\Delta_{p}(C_i, Q_i, F_i, s)/\sigma}},
\end{equation}
where $\sigma$ was a tuning parameter that measured how important it was to be near the optimal choice. A large $\sigma$ resulted in uniform probabilities of choices, with vessels being distributed uniformly across the different fishing areas. In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$). For computations, we used $\sigma = 3000$ (Table \ref{t:fishery}). %(noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).

The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota. Further details for this procedure can be found in \cite{Alzorriz2018, Batsleer2016} and \cite{Dowling2011}.

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}. For each year, these forward Monte Carlo simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area. The effort allocation component of the operating model provides the link between the management decisions and the biological component of the operating model.

\begin{table}[!h]
\centering
\caption{Model parameters.}
\label{t:fishery}
\resizebox{0.6\textwidth}{!}{%
\begin{tabular}{@{}llr@{}} 
\hline

\textbf{Population dynamics}                                    &                \\ \hline
&  \\%& & \\

number of recruits                                          & $R_i (p)$           & 500               \\
maximum age                                                 & $A$                 & 6               \\
number of areas                                             & $p$                 & 2               \\
number of months                                            & $S$                 & 12               \\
natural mortality                                           & $M_i$               & 1\e{-4} \\ %0.0001     \\
Asymptotic length                                           & $L_{\infty_i}$      & 50            \\
Growth rate                                                 & $K_i$               & 0.6         \\
Length-weight conversion factor                             & $ \alpha_i$         & 2\e{-4} \\% 0.0002   \\
Length-weight isomorphy factor                              & $\beta_i $          &   3              \\
Migration                                                   & $D_i $              & 2.5\e{-2} \\%  2.5\%              \\
&               \\
\textbf{Fleet dynamics}                                     &               \\ \hline   
&  \\%& & \\

Number of vessels                                           &                       & 8000           \\ 
Fuel costs (Euro fishing month\textsuperscript{-1})         &                       & 2000        \\ 
Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
Crew share                                                  &                       & 0\%        \\ 
Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
Optimal choice error                                        & $\sigma $             & 3000            \\

&                 \\  
\textbf{Fishery }                                           &             \\ \hline   
&  \\%& & \\

Intial quota  (kg vessel\textsuperscript{-1})               &                       & 200 \\%& & \\
catchability                                                & $q_i (a)$             & 2.5\e{-5} \\%0.000025        \\ 
effort (p,s)                                                &                       & 1                 \\ 
Price of species at mean weight (Euro kg\textsuperscript{-1})& $\bar{p}_i$          & 30             \\
Slope of species price                                      & $\gamma_i$            & 15            \\
Fine for overshooting quota  (Euro kg\textsuperscript{-1})  &  $F_i$                & 3000              \\
Ratio of standard deviations to catch means                 & $\eta$                & 0.08           \\ 
&                 \\  \hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependent pricing}

Prices of fish were assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}. Following \cite{Zimmermann2011}, therelationship between fish price and body weight was modelled as:  
\begin{equation}
 p_i (a,s) = \bar{p}_i + \gamma_i \times \frac{w_i (a,s) -\bar{w_i}}{\bar{w_i}},
\end{equation}
where $p_i (a,s) $ is the price of species $i$ at age $a$ and season $s$. The price was a linear function of the weight of individuals of species $i$ at age $a$ and season $s$. $\bar{p}_i$ is the price of the species at the mean weight over the age range. The mean weight over the age range is $\bar{w_i}$. $\gamma_i$ gave the price increase when individual mass was increased by $\bar{w_i}$.

\subsection{Case study parametrization}

To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery} and Fig. \ref{f:distribution}). There were two fish species in the model, both of which are caught by the fishery. The annual number of recruits was equal for the two species, and arbitrarily set to 500 per year (Table \ref{t:fishery}). The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Fig. \ref{f:distribution}). The maximum age that any individual can reach was 6 years old. During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Table \ref{t:fishery}). Conversion from length to weight was also equal for the two species (Fig. \ref{f:prices}). Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance (Table \ref{t:fishery}). For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Fig. \ref{f:distribution}). Mortality from natural causes such as predation, diseases, and senescence was assumed to be negligible, and $M_i$ set to 0.0001. Although such absence of natural mortality impossible in reality, one could argue that the model thus mimicked long-lived species.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Distributions.eps} 
\caption{Biomass distribution in numbers over the areas as a function of age when stocks are in virgin stock status for species 1 (a) and species 2 (b). For species 1 nursery ground occurs in the northern area, while species 2 in the southern area (high biomass in white). Over the two areas, initial biomass distribution shows a clear segregation of the younger ages (high and low fish biomass, in white and black respectively), while the older ages show similar distributions (grey colours).}
\label{f:distribution}
\end{figure}

Age-dependent catchability linked the population biomasses to the catches in the fishery, and was thus one of the crucial parameters determining the interaction between the two. In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species (Table \ref{t:fishery}).

The mean prices for the two species were set to 30 euro per kg, ranging between 17.3 euro per kg for the youngest age and 36.5  euro per kg for the oldest age (Table \ref{t:fishery} and Fig. \ref{f:prices}). The fines for overshooting the quota was set  to 3,000 euro per kg (Table \ref{t:fishery}). These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.

\begin{figure}[!ht]
\centering
\includegraphics[width=.5\textwidth]{Figures/Prices.eps} 
\caption{Lengths and prices of the two species.}
\label{f:prices}
\end{figure}

\subsection{Management procedure}

The management model encompassed the harvest control rules. The HCRs referenced to biological reference points ($H_{max}$) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort. The input to the fisheries model consisted of the expected catch rates and the individual quota that were set for each individual in the fleet. These individual quota were set in a management procedure, which mimicked the decisions of a management body. This management body made observations on the state of the resources, and the exploitation characteristics of the fishery. In the model, the management body was assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages. These observations stem from fisheries independent observations, such as surveys with known catchabilities. In addition, the management body made annual observations of the catches, including their age distribution.

The observed biomass and catches allowed for estimation of the annual harvest rates ($H_i$), and the estimation of a yield per recruit curve, that was dependent on fish growth and mortality. The yield curve was dome shaped, with a maximum that is called $H_{i,max}$. Fishing at a harvest rate that is equal to $H_{i,max}$  should lead to maximum sustainable yields. The yield per recruit curve omited the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment. The HCR used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$. These quotas were then divided equally over the individual vessels in the simulations.

\begin{equation}
 Q_{i,y+1} = \frac
                % Nominator
                {\sum( [\frac{H_{i,max}(a,y,s)}{\bar{H_i(a,s)}} \times H_i(a,s) \times sum( N_i (a, y, s, p))] \times w_i(a,s)) }
                % Denominator
                {number\  of\  vessels}.
\end{equation}

\subsection{Scenarios}

The model was set up in three consecutive time windows. In the absence of a plusgroup and 6 ages, the first window of 10 years was enough to get to stable populations. During this first period, new young individuals entered the population in the absence of fishing. This resulted in a virgin stock status. The consecutive 15 years, the fleet of 8000 vessels started fishing in the absence of any fisheries regulations. During this time window, the fishery was unmanaged, by meaned with unconstrained quota for both species (pre-manage period). During the last 15 years of projections, the management procedure was introduced, the fishery was constrained by setting MSY targets while discarding was not allowed (post-manage period).

Two scenarios based on two management scenarios were examined in the model. The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan. For this purpose, in a single-species catch quota context: quota on species 1 was was controlled by HCR in first scenario, while in both-species catch quota context: quotas on both species were controlled by HCRs in second scenario. The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY. To be able to analyse the results of this complex actions of implementing the MSY and the LO, we simulated the effects of implementing the LO without any exceptions or flexibility. When necessary, for example, in the case of estimating the modelled monthly mean of harvest rates, 5 years windows were selected to characterise the fleet in that pre- and post-manage period for both species (grey and green areas in Fig. \ref{f:catches}, respectively).

%changes in selectivity.  
%These management scenarios were chosen such that, if the management model was the correct model of reality, the corresponding equilibrium state of the system would meet the objective of the plan.

\section{Results}

% CATCHES
Stochastic simulations were carried out for the management scenarios contemplated: moving from an unmanaged period to a situation where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY, while it was not possible to discard.

Figure \ref{f:catches} shows the expected catch of the two target species and the impact of implementing the management scenarios on simulated catch. When the fishery was unmanaged, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$), thus resulted in a situation where the stocks were overfished, reaching total catches of more than 4 tons per species. In both scenarios, the introduction of PM caused an immediate reduction in catches of target species in the short-term. However, after such a reduction in the initial years, the stock biomass increases in the mid-term (after a transition of $<$5 years), and therefore mid-term TACs and catches were higher than before the implementation of the MP (Fig. \ref{f:catches}, c and d). The most notable differences in catch between the different scenarios lied in the absence of a regulatory quota for species 2. In the case where the fishery was limited only by HCR for species 1; higher catches for species 1 were observed in the mid-term. While species 2 maintained similar catches to those observed in the absence of management. As there was no quota restriction on species 2, catches of species 1 were adjusted to the estimated TAC in the HCR (Fig. \ref{f:catches}, a and b). However, for the fishery that was governed by quotas on both species, differences between catches and quotas were more noticeable due to the choke effect. As there was a similar spatial overlap of mature species and an antagonic spatial distribution of young species between the two fishing areas (Fig. \ref{f:distribution}), some fishing opportunities would always be lost for keeping the final fishing mortalities of both species below each target fishing mortality. The choke effect reduced the effort and made the fleet incapable of fishing their quota share.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Catches.eps} 
\caption{Modelled total annual catches (tonnes) for all vessels and both species: species 1 (a,c) and species 2 (b,d) in relation to the available individual quota (red line). In the top panels only species 1 quota constrained the fishery (a,b), while in the bottom panels both species quota constrained the fishery (c,d). MP year reflects the year where the management plan was introduced. Grey and green areas reflect the pre- and post-manage 5-year period, respectively.}
\label{f:catches}
\end{figure}

% HARVEST RATES
Modelling results indicated that the age composition of the harvest was different during the post-manage than it was during the pre-manage period (Fig. \ref{f:selectivity}). In the pre-manage period harvesting above the $H_{max}$ (with values around 0.03, Fig. \ref{f:catches}) caused instability in both populations. Overharvesting was caused by more significant harvesting in juvenile specimens than in adults, with a downward sloping harvest rate curves over fish age (rather than selectively targeting older fish with higher price; Fig.4), with average harvest rates of 0.06 (Fig. \ref{f:selectivity} and \ref{f:yields}, black lines). These harvests and levels of biomass remained stable over time due to the constant recruitment introduced into the population dynamics. On the other hand, after the MP implementation, there was a immediate decrease in harvest rates with values equal to or below $H_{max}$. After a few years, in the mid-term, 2 to 6-year old fish were harvested at the same rate during the post-manage period, while yearling fish (age class 1) was harvested at lower rates (Fig. \ref{f:selectivity} and \ref{f:yields}, grey lines).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Selectivity.eps} 
\caption{Modelled changes in harvest rates for both species in relation to the catch decision options made based in the availble individual quota: species 1 (a,c) and species 2 (b,d). In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d). Black lines: mean harvest rates during the pre-manage period (unconstrained fishery); grey line: mean harvest rates during the post-manage period (constrained fishery). Periods are relative to the introduction of the management plan (grey and green areas in Fig. \ref{f:catches}).}
\label{f:selectivity}
\end{figure}

% YIELDS 
Expected yields as a function of harvest show that during the post-manage period both species were underharvested to reach a sustainable harvesting (Fig. \ref{f:yields}; grey lines in a,c,d), except species 2 in the single-stock quota case scenario which was overharvested due to the abscence of any quota restriction (Fig. \ref{f:yields}; panel b). Indeed, higher sustainable yields achieved with lower values of fishing mortality. When quotas on both species, yields were 25\% higher than in the unmanaged situation; just on one specie quota, yield for specie 1 was 50\% higher than the unconstrined fishery value. These differences in quota scenarios could be explained due to the already commented choke effect. In general, the expected average yields shown similar values than the observed catches during the pre- and post-manage period selected; specially displaying little variability the MSY yields when accounting for quotas on both species scenario (Fig. \ref{f:yields}).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Yields.eps} 
\caption{Modelled yield per recruitment curves (tonnes) for both species in relation to the introduction of the management plan: species 1 (a,c) and species 2 (b,d). In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d). Black lines: expected yield during the pre-manage period (unconstrained fishery); grey line: expected yield during the post-manage period (constrained fishery). Periods are relative to the introduction of the management plan (grey and green areas in Fig. \ref{f:catches}). Blue dots represent the observed catches at the mean harvest rate for each year during the 5-year period (filled dots: pre-manage period and empty dots: post-manage period), while red dots are the expected yields at sustainable harvest rates.}
\label{f:yields}
\end{figure}

% EFFORT
The fishing effort that would lead the fishery toward the biological management reference points can be seen in Figure \ref{f:effort}. In an unmanaged fishing situation, given the spatial and price distribution of the species' parameters (Fig. \ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable. Therefore, 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 thousand Euros/ year (Fig. \ref{f:effort} b and d). Fuel costs related to fishing choices were 40\% of the incurred gross revenues (Fig. \ref{f:effort} b an d). Fishing in the northern (southern) area, during the pre-manage period, showed annual average catches of 3.1 tonnes/ year and 1.2 tonnes/year of species 1 and 2, respectively (1.2 tonnes/year and 3.1 tonnes/ year), which in terms of activity would be fishing around 3100 days per area and season (Fig. \ref{f:catchesbyarea} and  \ref{f:meaneffort} a and c). Some variability was appreciated between patch choices in the two scenarios (Fig. \ref{f:effort} a and c, \ref{f:catchesbyarea} and  \ref{f:meaneffort}); however, gross and net revenues at the end of each year resulted in similar values (Fig\ref{f:effort} b and d).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Efforteconomics.eps} 
\caption{Modelled spatial allocation of effort per year (\%) and the respective economic performance when only species 1 has quota limitations (a, b) and both species are quota limited (c,d). Trade-offs between net revenue (black line), gross revenue (grey line), fuel cost (blue line) and annual fines (red line) for the fleet are shown in panels (b,d). MP years reflect the year where the management plan was introduced. In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).}
\label{f:effort}
\end{figure}

When accounting in one species MSY quota, in the short-term, the fleet generated 42.31\% less revenue and a 37.50\% reduction in fishing effort than when compared to the unmanaged situation (Fig. \ref{f:effort}). At a low TAC of specie 1 in the short-term, the "effort-by-stock" required to catch species 1 TAC became smaller than the effort needed to harvest an unmanaged stock (species 2); thus, depending upon the different exploitation patterns of the fishing grounds. Therefore, vessels allocated 80\% (20\%) of the fishing effort in the southern (northern) fishing ground, while maintaining (reducing) both species catches (Fig. \ref{f:catchesbyarea}). On the other hand, such reductions in effort and changes in exploitation patterns led to an increase in stock biommass that generated a completely different picture in the mid-term. Maintaining the distribution of effort observed in the short-term, in the mid-term, the fleet would generate an average increase revenue of 76.92\% with respect to the pre-managed period, while average fuel cost were reduced to 20\% of the incurred gross revenues (Fig. \ref{f:effort} b). After the initial 57 \% to 32 \% decline in cathes of species 1 and 2 due to the MP implementation, fishing in the southern (northern) area, during the post-manage period, showed in average 181.67\% and 12.87\% higher catches of species 1 and 2 than the pre-manage period, respectively (16.91\% and 38.12\% lower catches of species 1 and 2)(Fig. \ref{f:catchesbyarea} a and b). In terms of activity the average number of fishing days per season droped to 3650 days (Fig. \ref{f:meaneffort} b). Vessels were incentivized to allocate the highest amount of effort early in the year (with around 1075 and 3700 fishing days in the northern and southern fishing grounds respectively), progressively decreasing to lower levels over the course of the year (with around 121 and 1634 fishing days per area; Fig. \ref{f:meaneffort} d).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Catchesbyarea.eps} 
\caption{Modelled catches (thousand tonnes/year) by area (blue line: Northern area, red line: Southern area) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period. In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery. MP year reflects the year where the management plan was introduced. Grey and green areas reflect the pre- and post-manage 5-year period, respectively.}
\label{f:catchesbyarea}
\end{figure}

When accounting in MSY quotas on both species, the inclusion of one more MSY quota revealed further reductions in fishing opportunities (Fig. \ref{f:effort}). In the short-term, the fleet generated 61.54\% less revenue and a 62.5\% reduction in fishing effort than when compared to the unmanaged situation. Vessels indistinctly allocated the fishing effort in both fishing grounds (such in the pre-manage period) due to same model conditioning for both species. Therefore, after the initial 62 \% to 65 \% decline in cathes due to the MP implementation, fishing in the northern (southern) area, during the post-manage period, somulations showed in average 2.57\% and 54.54\% higher catches of species 1 and 2 than the pre-manage period, respectively (55.46\% and 4.20\%)(Fig. \ref{f:catchesbyarea} c and d). The limiting species quota incentivized vessels to choose to allocate the highest amount of effort early in the year (with around 2500 fishing days per area), progressively decreasing to lower levels over the course of the year (with around 50 fishing days per area; Fig. \ref{f:meaneffort} d). In the mid-term, the fleet would generate an average increase revenue of 107.69\% with respect to the pre-managed period.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Meaneffort.eps} 
\caption{Modelled average spatial allocation of effort (days/month) during the pre-manage (a,c) and post-manage periods (b,d). Periods are relative to the introduction of the management plan (grey and green areas in Fig. \ref{f:catchesbyarea}).}
\label{f:meaneffort}
\end{figure}

Modelling results sugested that moving from an unmanaged period to a sustainable stock situation, where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY while it was not possible to discard, moderate and high reductions of fishing effort are expected (Fig. \ref{f:effort}, b and d). Thus, moderate to high reductions in the number of vessels modelled are expected too (effort cost per vessel 1 day; Fig. \ref{f:meaneffort}, b and d). The relative annual contribution by vessel to the gross revenue was arouond 28 Euros, reduced to 17 Euros of net revenues due to the costs associated to fuel during the unmanage situation (Fig. \ref{f:meanecon}). This implies that, in the short-term, revenues under the two scenarios are predicted to result in a loss of profitability (52.89\% when accounting in one species MSY quota and 34.43\% when accounting in both MSY quota; Fig. \ref{f:meanecon}). However, in the mid-term, when compared to the pre-manage period, accounting in one species MSY quota resulted in a 1.32 times larger gross revenues per vessel; while accounting in two species MSY quotas was 1.37 times larger benefits (Fig. \ref{f:meanecon}).

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{Figures/Mean_Economics.eps} 
\caption{Modelled main economics by vessel (euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively. MP year reflects the year where the management plan was introduced.}
\label{f:meanecon}
\end{figure}
%---

\section{Discussion}

Using a managent strategy evaluation framework, the potential consequences of two species quota scenarios were examined, allowing the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY. The MSE framework included temporal and spatial key processes in the dynamics of two fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995, Kell2007}. The analysis extended earlier approaches which used short-term state-dependent decision-making models (e.g. \cite{Alzorriz2018, Batsleer2016, Poos2010} by including the consequent biomass of fish stocks and the management procedure to make its decision on how to respond to the state of the resource. If the landing obligation changes selectivity patterns as predicted in \cite{Alzorriz2018, Batsleer2016}, the use of the present MSE allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling the feedback between fleet and stock dynamics. Nevertheless, the framework used made several assumptions: that there was perfect knowledge of the stochastic nature of catch rates, full compliance of fishers with the discarding policy, lack of transferability of individual quotas, and lack of interference competition among fishers \cite{Alzorriz2018}. Regarding fish populations, also important assumptions were established: constant mortality and almost neligible, constant recruitment along the years, equal conversion length to weight for the two species, constant migration that lead to a clear segregation of the youngers, age-dependant catchability assumed independent of age and equal for the two species and prices were age dependant but fixed over time. Also, we used monthly time steps in the model to keep computation time low, while the time scale at which fishing trips occur may be much smaller. Finally, we did not include the existence of any landing obligation discard flexibilities (CFP Article 15, \cite{CFP2013}) or year transfer swaps. We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under the landing obligation. Despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery fleet and examine whether potential harvest strategies can achieve the management objectives.

When the fishery was unmanaged, with unconstrained quota for both species (pre-manage period), the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$), thus resulted in a situation where the stocks were overfished, reaching total catches of more than 4 tons per species (Fig. \ref{f:catches}), which in terms of activity would be fishing around 3100 days per area and season (Fig. \ref{f:catchesbyarea} and \ref{f:meaneffort} a and c). Given the spatial and price distribution of the species' parameters (Fig. \ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable. Our results suggested that 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 thousand Euros/ year (Fig. \ref{f:effort} b and d). Fuel costs related to fishing choices were 40\% of the incurred gross revenues (Fig. \ref{f:effort} b an d).

Moving from an unmanaged period to a situation where one or both species were regulated by quotas and forced to comply simulataneously with management objectives (i.e. exploiting at MSY and no discarding allowance) on the optimal effort allocation between fleets was analysed. For example, the underlying assumption of the two quota scenarios that fleets would switch targeted stocks to their relative abundance and management constraints was tested. Under both scenarios, our model results showed that simultaneously implementation of MSY and LO objectives led to a very restrictive fishing activity. Effort is constrained by some quota stock, as predicted in general terms by \cite{Ulrich2017}. Short-term reductions in effort would be needed while mid-term interactions between fleets and stocks that occur in mixed fisheries such as t

%Consequently, no comparisons could be made to verify the hypothesis stating that an effort regulation is more biologically robust as there are no incentives for discards and black landings. This hypothesis would be interesting to investigate in the model as, according to Faroese scientists, it seems to be one of the main advantages of this management system %(Jákupsstovu & Reinert, personal communication).

Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests. We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model. The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations. Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation. Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation. The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.

\begin{itemize}
\item{Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity, at least, in the short-term. Effort is constrained by some stock, as predicted, in general terms, by Ulrich et al. (2011).}
\item{Drastic reduction of effort would increase profits if MSY and LO were contemplated. However, no flexibilities or year transfer swaps were considered. Be carefull if applying flexibilities, because as Prellezo 2016 (the results obtained from applying any kind of exemption or flexibility are, simply, negative) said, any increased in mortality beyond the MSY would be negative.}
\item{once the status of the stock has been assessed relative to its MSY level, the calculation of a TAC comes automatically from the rules of the Management Procedure. Shifting areas toward complying TAC limitations produced changes in selectivity ant therefore, constant changes in MSY levels. (e.g. shifting selectivity toward smaller/younger fish will generally reduce MSY). if it is required that the TAC should correspond to the level where MSY can be obtained, it will be necessary to monitor continuously the stock abundance and to revise TACs on a regular basis.}
\item{The  MSY under two quota species may be below what was achieved in one quota species because overall selectivity has shifted to smaller fish and low harvest of older fish; the impact of this change in selectivity on estimates of MSY is clearly seen in the results from Fig. \ref{f:selectivity} and \ref{f:yields}. While  the  potential  MSY  has decreased over time the stock biomass required to produce this MSY has increased.}
\item{Catch quota management is a more important mechanism than the actual obligation to land. An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.}
\item{The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics (Catchpole and Riveiro Santos, 2015) . Data risks should not be overlooked (Hentati-Sundberg et al., 2014). Fish stock assessment provides the foundation for scientific advice to decision-makers on sustainable catch levels. A critical feature of stock assessment methodology is its dependence on accurate commercial data, which makes it susceptible to biased reporting (Walters and Maguire, 1996; Cotter et al., 2004; Bousquet et al., 2010). The reasons for misreporting and underreporting include e.g. IUU fishing (Agnew et al., 2009), discard obligation (Kelleher, 2005; Wielgus et al., 2010) and catches from recreational fisheries (Zeller et al., 2011). Such misreporting can be driven by fishing fleet overcapacity, resulting from restrictive TACs and/or effort regulations (Patterson, 1998; Beare et al., 2005; Dobby et al., 2008) that create economic incentives for non-compliance (Sumaila et al., 2006; Gallic and Cox, 2006; Osterblom et al., 2011).}

\end{itemize}

%When discarding was allowed, fishers did not avoid areas where catching undersized specimens was likely because there was no incentive to do so. Under all three implementations (scenarios) of the landing obligation, however, our model showed that the Basque trawling fleet was likely to move to areas with lower catch rates of quota species to avoid exceeding quota limitations. We showed that in the short term the landing obligation would likely result in a loss in profitability for the fleet. This loss in profitability was driven by the ``choke'' effect for species such as mackerel (scenario II and III), which was triggered by reductions in effort as a result of rapid quota exhaustion and by greater costs associated with fishing further from home ports (scenario IV).

%Our results suggest that the Basque fleet is limited by hake, horse mackerel and mackerel quota. In scenarios II and III, the choke effect of mackerel is so strong that the effect of other species limits (hake under scenario II, horse mackerel under scenario III) is negligible. Consequently, under both these scenarios our model predicts that the fleet will avoid using OTB gear and will reduce fishing effort in sub-area 7 and 8 because of their high mackerel catch rates (Fig. \ref{fig:effort_modelval}). Thus, fishers will likely visit sub-area 6 and 8 (using PTB gear). Revenues under these two scenarios are predicted to result in a loss of profitability (60\% less than the status quo; Table ~\ref{t:net_rev}). These findings indicate that trading quotas (at rates undertaken in the past) does not alleviate the choking effect of the landing obligation. Therefore, new ways of swapping quotas would be needed under the landing obligation to alleviate the choke effect of any of the quota limiting species. Fleets with high discards rates of certain species will be likely to have larger incentives to increase their quotas for those species than fleets with low discard rates.

%In scenario IV, the addition of inter-species quota-swapping flexibility to the landing obligation (e.g. 9\% of the horse mackerel catch quota could go to increase mackerel quota) substantially decreased the mackerel choke effect and increased both fishing opportunities and net revenues, when compared to scenarios II and III. Under this scenario, revenue was 30\% lower than when discarding was allowed (scenario I).

%Our results suggest that reductions in fishing opportunities under a landing obligation that are caused by choke species could be alleviated by the use of inter-species quota flexibility, and that there will be strong incentives to use this flexibility. Management bodies should ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation. Use of inter-species quota flexibility could cause serious impacts to stock abundance when quota from a high volume/low value species is transferred to a low volume/high value stock. Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries.

%The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance. With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2005, Batsleer2013}. In our model, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations. However, if detection rates are low and the short–term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.

%The landing obligation was implemented to incentivise individual businesses to improve selectivity and avoid unwanted catches. The landing obligation shifts the emphasis from prescribing the fishing gears that are permissible, to a results-based system where the onus is on the fishers to use fishing gear that will avoid unwanted catches. Therefore, fishers should be incentivised to develop alternative gear or adapt their fishing behaviour to mitigate the full impact of the landing obligation \cite{Alzorriz2016}. Discard ban policies may not necessarily reduce harvest rates but will change the underlying selectivity patterns of harvests \cite{Borges2016}. Our results reflect the short-term effects of a discard ban on selectivity patterns and net revenues implemented under three scenarios. If the landing obligation changes selectivity patterns as we predict, then we would also expect longer term changes in both fish populations and the economy of the fishery. Such analyses are the domain of bio-economic management strategy evaluations where the feedback between fleet and stock dynamics are modelled explicitly \cite{Andersen2010, Prellezo2016a, Punt2016}. In such models, effects of the changing selectivity of the fishing fleets on the reference points for sustainable exploitation can be evaluated.

%Although this may seem implausible too, a drastic reduction of effort would increase profits if MSY and LO were contemplated. However, no flexibilities or year transfer swaps were consider. Be carefull if applying flexibilities, because as Prellezo said, any increased in mortality beyond the MSY would be negative...
%prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative. Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future. This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.

%The largest yield (or catch) that can be taken from a species' stock over an indefinite period. Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely. Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small. At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate. At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals. Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity. At this point, there are no surplus individuals to be harvested and yield drops to zero. The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield

%There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983). This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource. 
%For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis. Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.

%As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing. It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere. Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g. Panayotou 1988).

%The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced. This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g. the tuna fishery for the canning industry.

%The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition. Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points. Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.

%In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points. In evaluating investment projects, including resource management, economists discount the future value of any commodity. Discount rates may be in the order of 10\%. In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested. Long-lived species with slow growth rates, such as whales, clearly fall into this category. The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).

\begin{notes}[Acknowledgements]
The content of this paper does not reflect the official opinion of the European Commission. Responsibility for the information and views expressed in this paper relies entirely with the authors.
\end{notes}

\bibliographystyle{plain}
\bibliography{mylib}

\end{document}